---
title: "PBMC analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PBMC analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## Experimental setup

Peripheral mononuclear blood cells (PMBCs) from a healthy human were freshly prepared using Ficoll density separation. A suspension of ~10,000 cells in 30 microliters was prepared.

Prior to capture, 5 microliters of a solution of polyadenylated DNA hairpin repair substrates was added, containing:

1. A "normal" DNA substrate
1. A:U
1. G:U
1. Ribonucleotide (rG)
1. C:Inosine
1. T:Inosine

Cells were emulsified and the reactions were incubated at 37 ˚C for 1 hour prior to continuing with a single-cell RNA sequencing prep.

```{r example_load, message=FALSE}
library(scrunchy)
library(SingleCellExperiment)

# load mRNA UMI data
exp <- load_csv(scrunchy_data("pbmc.gene.matrix.csv.gz"))

# load haircut UMI data
fxn <- load_csv(scrunchy_data("pbmc.haircut.matrix.csv.gz"))

fce <- create_fce(exp, fxn)

fce <- normalize_counts(fce)

var_genes <- get_var_genes(fce, n_genes = 5000)
fce <- calc_pca(fce, n_pcs = 20, genes = var_genes)
fce <- calc_umap(fce, n_dims = 6)
```

### DNA repair activity

```{r repair_plot}
features <- c(
  "Uracil_45",
  "GU_45",
  "riboG_44",
  "Abasic_45",
  "TI_45",
  "CI_45"
)

plot_cells(fce, features)
```

### mRNA expression

mRNA expression of genes catalying known repair events.

```{r gene_features}
library(tibble)
gene_features <- tribble(
  ~ensembl, ~common,
  "ENSG00000076248", "UNG",
  "ENSG00000129071", "MBD4",
  "ENSG00000172922", "RNASEH2",
  "ENSG00000100823", "APEX1",
  "ENSG00000103152",  "MPG"
)
knitr::kable(gene_features)
```

```{r gene_plot}
plot_cells(fce, gene_features$ensembl)
```

