---
title: "Example usage of the scrunchy package"
author: "Kent Riemondy"
date: "`r Sys.Date()`"
output: radix::radix_article
vignette: >
  %\VignetteIndexEntry{Basic Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

# Install the package

```{r install, eval = FALSE}
devtools::install_github("hesselberthlab/scrunchy")
```

# Create a Haircut object

```{r create_sce}
suppressMessages(library(scrunchy))
suppressMessages(library(SingleCellExperiment))

# load mRNA UMI data
mrna <- scrunchy:::load_csv(scrunchy:::scrunchy_data("mrna.csv.gz"))
mrna[1:5, 1:5]

# load haircut UMI data
hcut <- scrunchy:::load_csv(scrunchy:::scrunchy_data("haircut.csv.gz"))
hcut[1:5, 1:5]


hce <- create_haircut(mrna, hcut)
```

`hce` is an object of class MultiAssayExperiment, which can contain data from multiple modalities, and
importantly does not require that each modality has the same number of cells or features. 

```{r}
# rna stores RNA info
hce[["rna"]]
counts(hce[["rna"]])[1:5, 1:5]

# hcut stores functional data
hce[["hcut"]]
counts(hce[["hcut"]])[1:5, 1:5]

# row annotations contain info about each row in the matrix
rowData(hce[["hcut"]])

# column annotations contain info about each row in the matrix, 
# or about each unique cell present in any of the experiments
colData(hce)
```

# Plots

`plot_hairpin` produces a coverage plot across each hairpin in the data.

```{r hairpin_plot}
plot_hairpin(hce)
```

Specific adduct positions can be highlighted by including this information in the `hce` object.

```{r hairpin_w_adduct_pos}
# add in adduct information when building hce object
adducts <- data.frame(hairpin = c("Uracil", "riboG"),
                      adduct_position1 = c(1, 10),
                      adduct_position2 = c(10, 50),
                      stringsAsFactors = FALSE)

hce <- create_haircut(mrna, hcut, adduct_positions = adducts)

rowData(hce[["hcut"]])

plot_hairpin(hce)
```

```{r}
## subset which cells are used for plots
plot_hairpin(hce,
             cells_to_plot = sample(colnames(hce[["hcut"]]), 1))
```

