---
title: "Introduction to scrunchy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to scrunchy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r echo=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)
```

# Create a Functional Cell Experiment (fce) object

```{r create_fce, message=FALSE}
library(scrunchy)
library(SingleCellExperiment)

# load mRNA UMI data
mrna <- load_csv(scrunchy_data("mrna.csv.gz"))

# load haircut UMI data
hcut <- load_csv(scrunchy_data("haircut.csv.gz"))

fce <- create_fce(mrna, hcut)
```

`fce` is an object of class `MultiAssayExperiment`, which can contain data from multiple modalities and does not require that each modality has the same number of cells or features. 

```{r}
# sce stores RNA info
fce[["sce"]]
counts(fce[["sce"]])[1:5, 1:5]

# fsce stores functional data
fce[["fsce"]]
counts(fce[["fsce"]])[1:5, 1:5]

# row annotations contain info about each row in the matrix
rowData(fce[["fsce"]])

# column annotations contain info about each row in the matrix, 
# or about each unique cell present in any of the experiments
colData(fce)
```

# Plots

`plot_hairpin` produces a coverage plot across each hairpin in the data.

```{r hairpin_plot}
plot_hairpin(fce)
```

Specific adduct positions can be highlighted by including this information in the `fce` object.

```{r hairpin_w_adduct_pos}
# add in adduct information when building fce object
adducts <- data.frame(
  hairpin = c("Uracil", "riboG"),
  adduct_position1 = c(1, 10),
  adduct_position2 = c(10, 50),
  stringsAsFactors = FALSE
)

fce <- create_fce(mrna, hcut, adduct_positions = adducts)

rowData(fce[["fsce"]])

plot_hairpin(fce)
```

```{r}
## subset which cells are used for plots
plot_hairpin(
  fce,
  cells_to_plot = sample(colnames(fce[["fsce"]]), 1)
)
```

# Normalization

RNA data is log-normalized and scaled to total UMI count for each cell. Functional data is normalized using the center log ratio method used for CITE-seq analysis.

Normalized data is stored in the `logcounts` slot of each `SingleCellExperiment` object.

```{r}
# normalize rna and functional data
fce <- normalize_counts(fce)

logcounts(fce[["fsce"]])[1:5, 1:3]
```

# Feature selection and PCA

```{r}
# get variable genes
var_genes <- get_var_genes(fce, n_genes = 5000) 

# calc PCA and embed in 2D
fce <- calc_pca(fce, n_pcs = 20, genes = var_genes)
```

# Embed cells in 2D using UMAP or tSNE

```{r}
fce <- calc_umap(fce, n_dims = 6)

fce <- calc_tsne(fce, n_dims = 6)
```

Cell embeddings are stored in the `reducedDims` slot of the `SingleCellExperiment` and can be accessed as a named list.

```{r}
reducedDims(fce[["sce"]])

reducedDims(fce[["sce"]])$UMAP[1:4, ]
```

# Visualization

Cell features can be visulized with the `plot_cells` function which will plot gene/function assay values or cell metadata. 

```{r}
features <- c("Uracil_45", "riboG_44","ENSG00000076248", "ENSG00000172922")
 
plot_cells(fce, features)

plot_cells(fce, features, method = "TSNE")
```

# Clustering
  
Kmeans clustering is implemented at this time, but additional functionality will be added. 
  
```{r kmeans}
fce <- run_kmeans(fce, 2)

plot_feature(fce, "k_cluster")
```
